### 2024-11-27 analysis/lod/insilico_mosaic_lod.Rmd
Add folders and files to analysis folder

---
title: "In Silico LOD Detection Evaluation"
date: '`r Sys.Date()`'
output: 
    bookdown::html_document2:
        toc: true
        toc_float: true
        df_print: paged
        code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(here)
```

# Objective
Summarize results of in silico experiment to determine the limit of detection for 
mosaic variant detect using the strelka2 somatic variant caller.

# Methods
- Mixtures of 300 X Illumina reads
    - Chromosome 20 only
    - AFs 
        - 0.00 All HG003
        - 0.01 98:2
        - 0.05 90:10
        - 0.10 80:20
        - 0.25 50:50
        - 0.5 All HG002
- Variant calling using strelka2, performed using the strelka2 app on DNAnexus
- Benchmarking with happy (vcf post processing - TODO add script reference) happy run using pFDA app
- Characterizing results
- Manual Curation
- Nuances and potential issues with using strelka2 for mosaic variant detection.

# Results
```{r}
insilico_lod_df <- read_excel(here("analysis/insilico_lod/Mixture Summary SNPs 01072020.xlsx"),
                              .name_repair = "universal")

tidy_lod_df <- insilico_lod_df %>%
    filter(AF != -1) %>%
    select(-starts_with("SNP")) %>%
    rename(
        fnsnp_pass_count = FN.SNP.PASS,
        fnsnp_pass_pct =  FN.SNP.PASS..,
        fnsnp_all_count = FN.SNP.ALL,
        fnsnp_all_pct = FN.SNP.ALL..,
        recall_pass_val = PASS.Metric.Recall,
        recall_all_val = All.Metric.Recall
    ) %>%
    pivot_longer(
        cols = c(
            fnsnp_pass_count,
            fnsnp_all_count,
            fnsnp_pass_pct,
            fnsnp_all_pct,
            recall_pass_val,
            recall_all_val
        ),
        names_to = c("metric", "var_filter", "met_type"),
        names_sep = "_",
        values_to = "value"
    ) 
```

- AF 0.1 were consistently detected by strelka2 and most variants were detected with AF 0.05. 
- Will start with the assumption that we are able to detect mosaic variants in the RM DNA down to 10% AF using strelka2.
- Can use duplex seq data to validate candidate and putative mosaic variants then optimize the filter parameters is appropriate.
```{r}

ggplot(tidy_lod_df) + 
    geom_vline(aes(xintercept = 0.1), linetype = 1) +
    geom_vline(aes(xintercept = 0.05), linetype = 2) +
    geom_point(aes(x = AF, y = value, fill = var_filter), shape = 21) + 
    facet_wrap(~metric*met_type, ncol = 1, scales = "free_y") + 
    theme_bw()

```

# Session Information
## System Information
```{r}
sessioninfo::platform_info()
```

## Package Versions
```{r}
sessioninfo::package_info() %>% 
    filter(attached = TRUE) %>% 
    select(package, loadedversion, date, source) %>%
    knitr::kable(booktabs = TRUE, row.names = FALSE)
```