### 2025-03-31 analysis/manuscript_figures.Rmd
cleaning up figure Rmarkdown

---
title: "Mosaic_Benchmark_Figures"
author: "Ade, Camille, and Nate"
date: "`r Sys.Date()`"
output: 
    bookdown::html_document2:
        toc: true
        toc_float: true
        df_print: paged
        code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(patchwork)
library(viridis)
library(ggpubr)
library(ggupset)
```

## Load and Tidy Input Data

### Variant VAFs DF for Figure 2
```{r}
## Loading manually curated variants data
chrom_order <- paste0("chr", c(1:22, "X", "Y"))
curated_variants <- read_tsv(here("data", "igv_curated_variants_RM.tsv")) %>%
    arrange(AF) %>%
    mutate(
        index = row_number(),
        CHROM = factor(CHROM, levels = chrom_order)
    ) %>%
    arrange(CHROM, POS, REF, ALT)

curated_variants_longer <- pivot_longer(curated_variants,
    cols = c("BGI_mut_freq", "Element_mut_freq", "Pacbio_mut_freq", "Illumina_mut_freq"),
    names_to = "Method",
    values_to = "VAF"
) %>%
    mutate(Method = str_replace(Method, "_mut_freq", "")) %>%
    mutate(Method = gsub("Illumina", "Illumina 300X", Method)) %>%
    mutate(Method = gsub("BGI", "BGI 100X", Method)) %>%
    mutate(Method = gsub("Element", "Element 136X", Method)) %>%
    mutate(Method = gsub("Pacbio", "Pacbio 108X", Method))

common_variants <- read_tsv(here("data", "RM_VAF_fig_70cand_updated.tsv")) %>%
    arrange(AF) %>%
    mutate(index = row_number())

curated_variants_longer_pacbio_sep <- read_tsv(here("data", "curated_variants_longer_pacbio_sep.tsv"))
```

### In Silico LOD Data
```{r}
insilico_lod_df <- list.files(here("analysis/lod/strelka2_chrom20_mixtures/all_summaries"), full.names = TRUE) %>%
    set_names(., str_extract(., "(?<=af)[:digit:]*(?=_result)")) %>%
    map_dfr(read_csv, .id = "AF") %>%
    mutate(AF = as.numeric(AF))
```

## Figures

### Figure 2
```{r}
annot_df <- data.frame(
    Variant_Category = "**Excluded** from Benchmark",
    l_xmin = -0.25,
    l_xmax = 25,
    l_ymin = -0.03,
    l_ymax = 0.05,
    r_xmin = 113,
    r_xmax = 137,
    r_ymin = -0.03,
    r_ymax = 0.05,
    l_text = "Lack of 300X support",
    l_textx = 6,
    l_texty = 0.4,
    r_text = "Lack of long read support",
    r_textx = 109,
    r_texty = 0.4,
    l_arrowx = -1.5,
    l_arrowy = 0.36,
    l_arrowxend = 1,
    l_arrowyend = 0.02,
    r_arrowx = 110,
    r_arrowy = 0.36,
    r_arrowxend = 116,
    r_arrowyend = 0.03
)

ortho_vafs_df <- curated_variants_longer %>%
    mutate(Variant_Category = case_when(
        RM_KEEP_REMOVE == "KEEP" ~ "Mosaic Benchmark",
        RM_KEEP_REMOVE != "KEEP" ~ "Excluded from Benchmark",
        TRUE ~ "ERROR"
    )) %>%
    mutate(Variant_Category = gsub(
        "((Excluded|Mosaic Benchmark))",
        "**\\1**",
        Variant_Category
    ))

(curated_ortho_vafs_fig <- ggplot(
    curated_variants_longer,
    aes(x = index, y = VAF)
) +
    annotate("rect",
        xmin = -Inf, xmax = Inf,
        ymin = 0.05, ymax = 0.3,
        alpha = 0.4, fill = "gray"
    ) +
    geom_line(
        data = ortho_vafs_df,
        aes(group = index), linewidth = 0.8
    ) +
    geom_point(
        data = ortho_vafs_df,
        aes(fill = Method), size = 5, shape = 21, color = "grey40"
    ) +
    scale_color_viridis() +
    theme_bw() +
    xlab("Ordered by Strelka2 VAF") +
    ylab("Sequencing Method VAF") +
    theme(
        axis.text = element_text(size = 14),
        plot.title = element_text(size = 22),
        axis.title.x = element_text(size = 22),
        axis.title.y = element_text(size = 22),
        axis.text.x = element_blank(),
        legend.title = element_text(size = 19),
        legend.text = element_text(size = 19),
        legend.position = "bottom", legend.box = "vertical"
    ) +
    theme(plot.title = element_text(hjust = 0.5)) +
    facet_wrap(
        ~ factor(Variant_Category,
            levels = c("**Mosaic Benchmark**", "**Excluded** from Benchmark")
        ),
        ncol = 1
    ) +
    theme(
        strip.text = ggtext::element_markdown(size = 20),
        legend.text = ggtext::element_markdown(size = 19)
    ) +
    geom_text(data = annot_df, aes(x = r_textx, y = r_texty, label = r_text), size = 6) +
    geom_text(data = annot_df, aes(x = l_textx, y = l_texty, label = l_text), size = 6) +
    geom_segment(
        data = annot_df,
        aes(x = l_arrowx, y = l_arrowy, xend = l_arrowxend, yend = l_arrowyend),
        arrow = arrow(length = unit(0.3, "cm"), type = "closed")
    ) +
    geom_segment(
        data = annot_df,
        aes(x = r_arrowx, y = r_arrowy, xend = r_arrowxend, yend = r_arrowyend),
        arrow = arrow(length = unit(0.3, "cm"), type = "closed")
    )
)
```

```{r}
ggsave(
    filename = here("figures", "fig2_curated-variants.pdf"),
    plot = curated_ortho_vafs_fig,
    height = 6, width = 11
)
```

### Figure 3

#### Curated Variants VAF

```{r mosaic_VAF_plot, message = FALSE}
(curated_variants_hist <- curated_variants %>%
    mutate(combined_VAF = x_ci_no_Dup / n_ci_no_Dup) %>%
    ggplot(aes(
        x = combined_VAF,
        color = RM_CURATION,
        fill = RM_CURATION
    )) +
    geom_histogram(color = "black", position = "identity", bins = 40) +
    geom_vline(xintercept = c(0.05, 0.3), linetype = "dashed", color = "black") +
    scale_color_manual(values = c("#00BFC4", "#F8766D")) +
    scale_fill_manual(
        values = c("#00BFC4", "#F8766D"),
        labels = c("Mosaic Benchmark (85)", "Excluded from Benchmark (50)")
    ) +
    labs(
        x = "Variant Allele Frequency (VAF)",
        y = "Count"
    ) +
    theme_light() +
    theme(
        legend.title = element_blank(), legend.position = "bottom",
        strip.text = element_blank()
    ) +
    facet_wrap(~RM_CURATION,
        ncol = 2,
        labeller = labeller(RM_CURATION = c(
            "MOSAIC" = "Mosaic Benchmark",
            "NOT MOSAIC" = "Excluded from Benchmark"
        ))
    )
)
```

#### Genomic Context Upset

```{r mosaic_variant_upset_plot, message = FALSE}
# Create upset plot using GRCh38 genomic context annotations from curated variants df
(upset_plot <- curated_variants %>%
    mutate(
        lowmappability = if_else(easy_to_map == "hit", FALSE, TRUE),
        not_lowmappability = if_else(easy_to_map == "hit", TRUE, FALSE),
        homopolymer = if_else(homopolymer_region == "hit", TRUE, FALSE),
        nonhomopolymer = if_else(homopolymer_region == "hit", FALSE, TRUE)
    ) %>%
    group_by(CHROM, POS, REF, ALT, RM_CURATION) %>%
    pivot_longer(
        cols = c(not_lowmappability, lowmappability, homopolymer, nonhomopolymer),
        names_to = "genomic_context",
        values_to = "hit"
    ) %>%
    filter(hit) %>%
    mutate(RM_CURATION = factor(RM_CURATION, levels = c("NOT MOSAIC", "MOSAIC"))) %>%
    summarise(genomic_context = list(genomic_context)) %>%
    ggplot(aes(x = genomic_context, fill = RM_CURATION)) +
    geom_bar(color = "black") +
    scale_x_upset() +
    labs(x = "Genomic Context", y = "Count") +
    scale_fill_discrete(labels = c("Excluded from Benchmark", "Mosaic Benchmark")) +
    theme_bw() +
    theme(legend.title = element_blank(), legend.position = "none"))
```

#### Create panel containing VAF and upset plot

```{r histo_upset panel}
(histo_upset_panel <- curated_variants_hist /
    upset_plot +
    plot_annotation(tag_levels = "a") &
    theme(plot.tag = element_text(face = "bold")) &
    plot_layout(guides = "collect") +
        theme(legend.position = "none"))
```

```{r}
ggsave(
    plot = histo_upset_panel,
    filename = here("figures", "fig3_curated-variants.pdf"),
    height = 6, width = 11
)
```

### Figure 4

```{r}
# Bringing in the Non-RM data and selecting only Element and Pacbio results
non_RM_curated_variants <- curated_variants_longer_pacbio_sep %>%
    select(
        CHROM, POS, REF, ALT, AF,
        Element_depth, Element_alt_depth, Element_map_qual, Element_base_qual,
        Pacbio_revio_depth, Pacbio_revio_alt_depth, Pacbio_revio_map_qual, Pacbio_revio_base_qual,
        Pacbio_sequel_depth, Pacbio_sequel_alt_depth, Pacbio_sequel_map_qual, Pacbio_sequel_base_qual,
        Method, VAF
    ) %>%
    filter(Method %in% c("Element", "Pacbio_revio", "Pacbio_sequel")) %>%
    mutate(Source = "Non RM") %>%
    mutate(Element_mutfreq_nonRM = Element_alt_depth / Element_depth) %>%
    mutate(Pacbio_revio_mutfreq_nonRM = Pacbio_revio_alt_depth / Pacbio_revio_depth) %>%
    mutate(Pacbio_sequel_mutfreq_nonRM = Pacbio_sequel_alt_depth / Pacbio_sequel_depth) %>%
    select(
        CHROM, POS, REF, ALT, AF,
        Element_mutfreq_nonRM,
        Pacbio_revio_mutfreq_nonRM,
        Pacbio_sequel_mutfreq_nonRM
    )

common_variants_pacbio_element_sep <- pivot_longer(common_variants,
    cols = c(
        "BGI_mut_freq", "Element_standard_mut_freq", "Element_long_mut_freq",
        "Pacbio_revio_mut_freq", "Pacbio_sequel_mut_freq", "Illumina_mut_freq"
    ),
    names_to = "Method",
    values_to = "VAF"
) %>%
    mutate(Method = str_replace(Method, "_mut_freq", "")) %>%
    mutate(Method = gsub("Illumina", "Illumina 300X", Method))

rm_vs_nonrm_pairwise <- common_variants_pacbio_element_sep %>%
    select(
        CHROM, POS, REF, ALT, AF,
        Element_depth, Element_alt_depth, Element_map_qual, Element_base_qual,
        Element_standard_depth, Element_standard_alt_depth, Element_standard_map_qual, Element_standard_base_qual,
        Element_long_depth, Element_long_alt_depth, Element_long_map_qual, Element_long_base_qual,
        Pacbio_revio_depth, Pacbio_revio_alt_depth, Pacbio_revio_map_qual, Pacbio_revio_base_qual,
        Pacbio_sequel_depth, Pacbio_sequel_alt_depth, Pacbio_sequel_map_qual, Pacbio_sequel_base_qual,
        Method, VAF
    ) %>%
    filter(Method %in% c("Element", "Element_standard", "Element_long", "Pacbio_revio", "Pacbio_sequel")) %>%
    mutate(Source = "RM") %>%
    mutate(Element_standard_mutfreq_RM = Element_standard_alt_depth / Element_standard_depth) %>%
    mutate(Element_long_mutfreq_RM = Element_long_alt_depth / Element_long_depth) %>%
    mutate(Element_mutfreq_RM = Element_alt_depth / Element_depth) %>%
    mutate(Pacbio_revio_mutfreq_RM = Pacbio_revio_alt_depth / Pacbio_revio_depth) %>%
    mutate(Pacbio_sequel_mutfreq_RM = Pacbio_sequel_alt_depth / Pacbio_sequel_depth) %>%
    select(
        CHROM, POS, REF, ALT, AF,
        Element_mutfreq_RM, Element_standard_mutfreq_RM, Element_long_mutfreq_RM,
        Pacbio_revio_mutfreq_RM, Pacbio_sequel_mutfreq_RM
    ) %>%
    left_join(
        non_RM_curated_variants,
        by = c("CHROM", "POS", "REF", "ALT")
    ) %>%
    select(-AF.x) %>%
    rename(AF = AF.y) %>%
    distinct() %>%
    arrange(AF) %>%
    mutate(index = row_number())

rm_vs_nonrm_pairwise_preprint <- rm_vs_nonrm_pairwise %>%
    select(Element_mutfreq_RM, Element_mutfreq_nonRM) %>%
    rename(mutfreq_RM = Element_mutfreq_RM, mutfreq_nonRM = Element_mutfreq_nonRM) %>%
    mutate(Method = "Element") %>%
    rbind(rm_vs_nonrm_pairwise %>%
        select(Pacbio_revio_mutfreq_RM, Pacbio_revio_mutfreq_nonRM) %>%
        rename(mutfreq_RM = Pacbio_revio_mutfreq_RM, mutfreq_nonRM = Pacbio_revio_mutfreq_nonRM) %>%
        mutate(Method = "Pacbio Revio"))

rm_vs_nonrm_pairwise_preprint_Element <- rm_vs_nonrm_pairwise_preprint %>%
    filter(Method == "Element") %>%
    mutate(mutation = row_number())

rm_vs_nonrm_pairwise_preprint_Pacbio_Revio <- rm_vs_nonrm_pairwise_preprint %>%
    filter(Method == "Pacbio Revio") %>%
    mutate(mutation = row_number())

wilcox_rm_vs_nonrm_pairwise_preprint_Element <- rm_vs_nonrm_pairwise_preprint_Element %>%
    rename("non-RM" = mutfreq_nonRM, "RM" = mutfreq_RM) %>%
    pivot_longer(cols = c("non-RM", "RM"), names_to = "Sample Source", values_to = "VAF")

wilcox_rm_vs_nonrm_pairwise_preprint_Pacbio_Revio <- rm_vs_nonrm_pairwise_preprint_Pacbio_Revio %>%
    rename("non-RM" = mutfreq_nonRM, "RM" = mutfreq_RM) %>%
    pivot_longer(
        cols = c("non-RM", "RM"),
        names_to = "Sample Source",
        values_to = "VAF"
    )

wilcox_rm_vs_nonrm_pairwise_preprint <- rbind(wilcox_rm_vs_nonrm_pairwise_preprint_Element, wilcox_rm_vs_nonrm_pairwise_preprint_Pacbio_Revio)
```

```{r}
(RMvsNonRM_paired_wilcox_preprint <- ggpaired(wilcox_rm_vs_nonrm_pairwise_preprint,
    x = "Sample Source", y = "VAF", id = "mutation",
    color = "Method", line.color = "gray", line.size = 0.4,
    palette = c("#7CAE00", "#C77CFF"),
    facet.by = "Method"
) +
    stat_compare_means(paired = TRUE, size = 6) +
    ylim(0, 0.4) +
    ylab("Variant Allele Frequency (VAF)") +
    xlab("Sample Source") +
    theme(
        plot.title = element_text(hjust = 0.5, size = 22),
        legend.position = "none",
        axis.text = element_text(size = 14),
        axis.title.x = element_text(size = 22),
        axis.title.y = element_text(size = 22),
        strip.text = element_text(size = 16),
        axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 16)
    ))
```

```{r}
ggsave(
    plot = RMvsNonRM_paired_wilcox_preprint,
    filename = here("figures", "fig4_VAF-RM-v-nonRM.pdf")
)
```

### Figure S1

```{r}
(lod_fig <- insilico_lod_df %>%
    ggplot() +
    geom_vline(aes(xintercept = 10), linetype = 1) +
    geom_vline(aes(xintercept = 5), linetype = 2) +
    geom_path(aes(x = AF, y = METRIC.Recall, color = Filter, linetype = Type)) +
    geom_point(aes(x = AF, y = METRIC.Recall, fill = Filter, shape = Type)) +
    scale_shape_manual(values = c(21, 22)) +
    theme_bw() +
    labs(x = "Allele Frequency (%)", y = "Recall") +
    theme(legend.position = "bottom"))
```

```{r}
ggsave(
    plot = lod_fig,
    filename = here("figures", "figS1_insilico-lod.pdf")
)
```

### Figure S4
```{r}
(curated_vafs_hist <- ggplot(curated_variants_longer, aes(x = VAF, fill = Method)) +
    geom_histogram(color = "black", alpha = 0.7) +
    facet_wrap(Method ~ ., scales = "fixed", ncol = 1) +
    scale_fill_manual("Method",
        breaks = c("BGI 100X", "Element 136X", "Illumina 300X", "Pacbio 108X"),
        values = c("#F9746D", "#7CAE00", "#00BEC4", "#C77CFF")
    ) +
    theme_bw() +
    theme(legend.position = "none") +
    xlab("Variant Allele Frequency (VAF)") +
    ylab("Number of Manually Curated Variants"))
```

```{r}
ggsave(
    plot = curated_vafs_hist,
    filename = here("figures", "figS4_curated-VAFs.pdf"),
    height = 8, width = 6
)
```

### Figure S5
```{r}
(bench_vafs_hist <- curated_variants_longer %>%
    filter(RM_KEEP_REMOVE == "KEEP") %>%
    ggplot(aes(x = VAF, fill = Method)) +
    geom_histogram(color = "black", alpha = 0.7) +
    facet_wrap(Method ~ ., scales = "fixed", ncol = 1) +
    scale_fill_manual("Method",
        breaks = c("BGI 100X", "Element 136X", "Illumina 300X", "Pacbio 108X"),
        values = c("#F9746D", "#7CAE00", "#00BEC4", "#C77CFF")
    ) +
    theme_bw() +
    theme(legend.position = "none") +
    xlab("Variant Allele Frequency (VAF)") +
    ylab("Number of Mosaic Variants"))
```

```{r}
ggsave(
    plot = bench_vafs_hist,
    filename = here("figures", "figS5_benchmark-VAFs.pdf"),
    height = 8, width = 6
)
```

# Session Information
## System Information

```{r}
sessioninfo::platform_info()
```

## Package Versions
```{r}
sessioninfo::package_info() %>%
    filter(attached = TRUE) %>%
    select(package, loadedversion, date, source) %>%
    knitr::kable(booktabs = TRUE, row.names = FALSE)
```